Bootstrap learning via modular concept discovery

### 2013-04-22 drafts/ijcai2013/camera_ready/Dechter.IJCAI13.CameraReady.tex
Inserted the rest of Josh's edits.

This structure derives from domain specific knowledge. In local
search, domain knowledge provides the similarity metric between
solutions; in heuristic search it provides the heuristic. Here we ask
how to bootstrap domain specific knowledge and extract structure from
the space of possible solutions by attempting to solve multiple
related problems in a given domain at once. Our approach is motivated
by the observation that many interesting problems -- such as
discovering scientific theories, designing new technologies, learning
a skill such as playing an instrument or cooking -- have solutions that
are naturally specified as programs.

@@

A natural metric for the compressibility of a set of binary trees is
the number of unique subtrees (counting leaves) that this set
contains. More sophisticated metrics exist; we could use the negative
logarithm of the probability of the solutions under a suitable
probability distribution~\cite{DBLP:journals/tit/BarronRY98}. Future
work will explore whether this sophistication is worthwhile and which
prior on grammars makes sense. For our purposes here, penalizing
according to the number of unique subtrees appropriately rewards reuse
of useful subtrees. As an example of this, suppose we have six tasks
whose solutions are integers $4^2, 5^2, 6^2, 7^2$, and $8^2$, and
suppose our frontier contains the $10$ expressions:
\vspace{-.3cm}
\captionsetup[subfigure]{labelformat=empty} 
\begin{figure}[h!]
   \begin{subfigure}{.17\linewidth}
     \begin{tikzpicture}[scale=.7, background rectangle/.style={fill=lightgray}] 
       \tikzset{level distance=15pt, sibling distance=10pt}
       \Tree [[  * 4 ] 4 ]
     \end{tikzpicture}
  \end{subfigure}
  \begin{subfigure}{.17\linewidth}
     \begin{tikzpicture}[scale=.7]
       \tikzset{level distance=15pt, sibling distance=10pt}
       \Tree [[  * 5 ] 5 ]
     \end{tikzpicture}
  \end{subfigure}
   \begin{subfigure}{.17\linewidth}
     \begin{tikzpicture} [scale=.7]
       \tikzset{level distance=15pt, sibling distance=10pt}
       \Tree [[  * 6 ] 6 ]
     \end{tikzpicture}
  \end{subfigure}
   \begin{subfigure}{.17\linewidth}
     \begin{tikzpicture}[scale=.7] 
       \tikzset{level distance=15pt, sibling distance=10pt}
       \Tree [[  * 7 ] 7 ]
     \end{tikzpicture}
  \end{subfigure}
  \begin{subfigure}{.17\linewidth}
     \begin{tikzpicture}[scale=.7] 
       \tikzset{level distance=15pt, sibling distance=10pt}
       \Tree [[  * 8 ] 8 ]
     \end{tikzpicture}
  \end{subfigure}
  \vspace{2mm}

@@

@@

In Figure~\ref{fig:symregCurricula}, we present learning curves
(frontier size 10000) corresponding to various versions of our
original symbolic regression task set. Recall that the original set
consisted of all polynomials of degree two or less and with
coefficients between 0 and 9. In the ``no constant'' task set, we
remove the 9 constant functions in this set. In the ``no linear'' task
set, we remove 90 linear functions from the set. We observe that
performance on those task sets does not decrease. However, when we
remove both the linear functions and the constant function (``only
quadratics''), we see a sharp drop in the algorithm's
performance. When we further restrict the task set to only ``complex''
quadratics (which we define as quadratics whose
coefficients are greater than zero), we observe another comparable
drop in performance. When we go one step further and restrict the task
set to quadratics with coefficients greater than 1, performance drops
to 0 because no task is hit in the initial frontier.

@@

The problem presented here was chosen to highlight the vastness of the
search space confronting an agent with minimal background knowledge:
the reward function is binary, it is deterministic, and it cannot
provide feedback about partial solutions. An important question left
to be addressed is how to extend this algorithm to take into account
continuous rewards, noisy data, and partial solutions.

>>>>

This structure derives from domain specific knowledge. In local
search, domain knowledge provides the similarity metric between
solutions; in heuristic search it provides the heuristic. Here we ask
how to bootstrap domain specific knowledge and extract structure from
the space of possible solutions by attempting to solve multiple
related problems in a given domain at once. Our approach is motivated
by the observation that many interesting problems -- such as
discovering scientific theories, designing new technologies, learning
a new language or learning a skill such as playing an instrument or
cooking -- have solutions that are naturally specified as programs.

@@

A natural metric for the compressibility of a set of binary trees is
the number of unique subtrees (counting leaves) that this set
contains. As an example of this, suppose we have six tasks
whose solutions are integers $4^2, 5^2, 6^2, 7^2$, and $8^2$, and
suppose our frontier contains the $10$ expressions:
\vspace{-.3cm}
\captionsetup[subfigure]{labelformat=empty} 
\begin{figure}[h!]
   \begin{subfigure}{.17\linewidth}
     \begin{tikzpicture}[scale=.7, background rectangle/.style={fill=lightgray}] 
       \tikzset{level distance=15pt, sibling distance=10pt}
       \Tree [[  * 4 ] 4 ]
     \end{tikzpicture}
  \end{subfigure}
  \begin{subfigure}{.17\linewidth}
     \begin{tikzpicture}[scale=.7]
       \tikzset{level distance=15pt, sibling distance=10pt}
       \Tree [[  * 5 ] 5 ]
     \end{tikzpicture}
  \end{subfigure}
   \begin{subfigure}{.17\linewidth}
     \begin{tikzpicture} [scale=.7]
       \tikzset{level distance=15pt, sibling distance=10pt}
       \Tree [[  * 6 ] 6 ]
     \end{tikzpicture}
  \end{subfigure}
   \begin{subfigure}{.17\linewidth}
     \begin{tikzpicture}[scale=.7] 
       \tikzset{level distance=15pt, sibling distance=10pt}
       \Tree [[  * 7 ] 7 ]
     \end{tikzpicture}
  \end{subfigure}
  \begin{subfigure}{.17\linewidth}
     \begin{tikzpicture}[scale=.7] 
       \tikzset{level distance=15pt, sibling distance=10pt}
       \Tree [[  * 8 ] 8 ]
     \end{tikzpicture}
  \end{subfigure}
  \vspace{2mm}

@@

More sophisticated metrics than the one proposed here exist; we could
use the negative logarithm of the probability of the solutions under a
suitable probability
distribution~\cite{DBLP:journals/tit/BarronRY98}. Future work will
explore whether this sophistication is worthwhile and which prior on
grammars makes sense. For our purposes here, penalizing according to
the number of unique subtrees appropriately and sufficiently rewards
reuse of useful subtrees.

@@

In Figure~\ref{fig:symregCurricula}, we present learning curves
(frontier size~10000) corresponding to various versions of our
original symbolic regression task set. Recall that the original set
consisted of all polynomials of degree two or less and with
coefficients between 0 and 9. In the ``no constant'' task set, we
remove the 9 constant functions in this set. In the ``no linear'' task
set, we remove 90 linear functions from the set. We observe that
performance on those task sets does not decrease. However, when we
remove both the linear functions and the constant function (``only
quadratics''), we see a sharp drop in the algorithm's
performance. When we further restrict the task set to only ``complex''
quadratics (which we define as quadratics whose
coefficients are greater than zero), we observe another comparable
drop in performance. When we go one step further and restrict the task
set to quadratics with coefficients greater than 1, performance drops
to 0 because no task is hit in the initial frontier.

@@

The problem presented here was chosen to highlight the vastness of the
search space confronting an agent with minimal background knowledge:
the reward function is binary, it is deterministic, and it cannot
provide feedback about partial solutions. In ongoing work we are
exploring a more probabilistic version of the algorithm that can take
into account continuous rewards, noisy data, and partial solutions.
